Unprepared colonoscopy identifies microbiota specific to the proximal and distal human colon

### 2017-07-18 submission/manuscript.tex
more edits

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Specific taxa differentiate proximal and distal microbiota in the unprepped human colon},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{Specific taxa differentiate proximal and distal microbiota in the
unprepped human colon}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

@@

Running title: Specific taxa differentiate proximal and distal human
colon microbiota

@@

Colorectal cancer (CRC) remains a leading cause of death worldwide
despite improved preventative and therapeutic measures. Tumors of the
proximal (proximal) and distal (distal) colon are morphologically and
genetically distinct. Previous work from our group found that microbial
dysbiosis is associated with the development of CRC tumors in studies of
both mice and humans. Analysis of the fecal microbiota from healthy and
CRC patients further revealed different microbial signatures associated
with disease. We extended our observations of the fecal microbiome to
include analysis of the proximal and distal human colon. We used a
two-colonoscope approach on subjects that had not undergone standard
bowel preparation procedure. This technique allowed us to characterize
the native proximal and distal luminal and mucosal microbiome without
prior chemical disruption. 16S rRNA gene sequencing was performed on
proximal and distal mucosal and luminal biopsies and home-collected
stool for 20 healthy individuals. Diversity analysis revealed that each
site contained a diverse community, and that a patient's samples were
more similar to each other than to that of other individuals. Since we
could not differentiate sites along the colon based on community
structure or community membership alone, we employed the Random Forest
machine-learning algorithm to identify key species that distinguish
biogeographical sites. Random Forest classification models were built
using taxa abundance and sample location and revealed distinct
populations that were found in each location. \emph{Peptoniphilus,
Anaerococcus, Enterobacteraceae, Pseudomonas} and \emph{Actinomyces}
were most likely to be found in mucosal samples versus luminal samples
(AUC = 0.925). The classification model performed well (AUC = 0.912)
when classifying mucosal samples into proximal or distal sides, but
separating luminal samples from each side proved more challenging (AUC =
0.755). The distal mucosa was found to have high populations of
\emph{Finegoldia, Murdochiella} and \emph{Porphyromonas}. Proximal and
distal luminal samples were comprised of many of the same taxa, likely
reflecting the fact that stool moves along the colon from the proximal
to distal end. Finally, comparison of all samples to fecal samples taken
at exit uncovered that the feces were most similar to samples taken from
the distal lumen, again reflecting the anatomical structure of the
colon. By sampling the unprepped human colon, our results have
identified distinct bacterial populations native to the proximal and
distal sides. Further investigation of these bacteria may elucidate if
and how these groups contribute to differential oncogenesis processes on
the respective sides of the colon.

@@

Colorectal cancer (CRC) is the second-leading cause of cancer-related
deaths in the United States. CRC tumors vary in structure, size,
morbidity and symptomology depending upon their geographical location in
the colon. CRC has recently been described to fall into four molecular
subtypes ((1)). Of these subtypes, tumors that arise on the distal side
of the colon are infiltrating lesions that present with painful
symptoms. In contrast, 47\% of CRCs are caused by proximal-sided colon
tumors that are bulky and project into the lumen, often remaining
asymptomatic until advanced carcinogenesis ((2)). Due to the absence of
symptoms, proximal-sided tumors have a significantly higher mortality
rate ((2)). The distal and proximal sides of the colon differ in the
amount of inflammation present and the genomic instability of
precancerous cells, respectively, as well as oxygen, pH and the presence
of antimicrobial peptides ((3), (4), (5)). Microenvironments differ not
only longitutinally along the colon, but latitudinally from the
epithelium to mucosa to intestinal lumen, offering several sites for
different microbial communities to flourish. Given this varied
physiology of the proximal-distal axis of the colon, symbiotic microbes
and their metabolites likely vary between sites.

Several recent findings have shown that development and progression of
CRC can be attributed to specific molecular events as a result of
interactions between the gut microbiota and human host ((3)). Our group
and others have found that the stool microbiome of patients with CRC is
distinct from that of healthy people ((6)). Further studies manipulating
the gut microbiome using antibiotics or other chemoagents in mice has
shown that dysbiosis preceded and accelerated the development of CRC
tumors ((7)). Comparison of the specific bacteria present on CRC tumors
with those found on nearby healthy tissue has also identified specific
bacterial species that are tumor-associated ((8)). These species include
the oral pathogens \emph{Fusobacterium nucleatum} and
\emph{Porphyromonas asacharolytica}. Interestingly, these periodontal
pathogens have been highly predictive of whether a patient had CRC
tumors or not in our classification studies ((9)). Because these studies
were performed in mice or examined only shed human stool, they were
unable to analyze paried samples from the proximal and distal sides of
the colon. Similarly, comparisons of on- or off-tumor bacteria rarely
have matched tissue from the other side of the colon from the same
patient, limiting what conclusions can be drawn about the colonic
microbiome overall, let alone at that specific site. Due to these
limitations, the contribution of the gut microbiota to these subtypes is
largely undefined. Characterizing these communities could provide needed
insight into CRC etiology, including how the disruption of the healthy
community could promote the initiation or proliferation of the distinct
distal and proximal CRC tumors.

Further, the few existing profiles of the microbial biogeography of the
gut have been limited by sample collection methods. The majority of
human gut microbiome studies have been performed on whole shed stool or
on samples collected during colonoscopy procedures. While the latter
method allows investigators to acquire samples from inside the human
colon, typically this procedure is preceded by the use of bowel
preparation methods such as the consumption of laxatives to cleanse the
bowel. Bowel preparation is essential for detecting cancerous or
precancerous lesions in the colon, but complicates microbiome profiling
as the chemicals strip the bowel of contents and disrupt the mucosal
layer ((10), (11)). As such, what little information we do have about
the biogeographical distribution of the microbes in the proximal and
distal colon is confounded by the bowel preparation procedure.

Here we aimed to address the limitations of previous studies and
effectively characterize the microbiome in the lumen and mucosa of the
proximal and distal healthy human colon. Our design used unprepared
colonoscopy techniques to sample the natural community of each location
of the gut without prior disruption of the native bacteria in 20 healthy
volunteers. To address the inherent inter-individual variation in human
microbiomes, we used a machine-learning classification algorithm trained
on curated 16S rRNA sequencing reads to identify microbes specific to
each location. We found that our classification models were able to
separate mucosal and lumenal samples as well as differentiate between
sides of the colon based on populations of specific microbes. By
identifying the specific microbes we are poised to ask if and how the
presence or disruption of the microbes at each site contribute to the
development of the specific tumor subtypes of CRC in the proximal and
distal human colon.

@@

Lumenal and mucosal samples were collected from the proximal and distal
colon of 20 healthy humans that had not undergone bowel preparation
(Figure 1). Participants also collected stool at home one week prior to
the procedure. To characterize the bacterial communities present at
these sites, 16S rRNA gene sequencing was performed on extracted DNA
from each sample. Each site was primarily dominated by \emph{Firmicutes}
and \emph{Bacteriodetes} (Figure 2A), consistent with known variability
in human microbiome research (cite). Likewise, samples had varying
levels of diversity at each site, irrespective of the individual (Figure
2B). For example, the proximal mucosa was more diverse than the distal
for some individuals while the opposite was true for others. Therefore
we could not identify a clear pattern of changes in microbial diversity
along the gut axis.

@@

Here we identified bacterial taxa that are specific to the lumen and
mucosa of the proximal and distal human colon from samples collected
during unprepared colonoscopy. We found that all locations contained a
range of phyla and a range of diversity, but that there was a wide
variability between subjects. Pairwise comparisons of each of the sites
revealed that the proximal mucosa and lumen were most similar to each
other. Further, comparison of colonoscopy-collected samples with samples
collected from stool at home showed that the distal lumen is most
similar to stool at exit. Random Forest algorithms built on OTU
abundances from each sample identified microbes that are particular to
each location of the colon. Finally, we were able to detect some
bacterial OTUs associated with colorectal cancer in our healthy patient
cohort. Using unprepped colonoscopy and machine learning, we have
identified bacterial phyla specific to the healthy proximal and distal
human colon.

When examining the relative abundance of the different phyla at each
site, there was a wide amount of variation for each phyla with
communities primarily dominated by the Bacteriodes and Firmicutes. This
likely reflects not only the variability between human subjects, casued
by differences in age, gender, diet, but also reports of microbial
``patchiness'' in the gut microbiome. Several previous studies have
noted that the bacteria recoverable from the same mucosal sample
location can be vastly different when the samples are taken just 1 cm
away from each other ((15)). Similar patchiness is also observed in
lumenal contents and fecal samples themselves; there is observed
separation of different interacting microbes along the length of a stool
sample, for instance ((16)). That said, across our samples the mucosal
samples harbor more Proteobacteria, consistent with previous studies
comparing mucosal swabs to lumenal content in humans ((4)). Hence, the
conclusions we can draw from phyla analysis are likely confounded by
differences in sampling and patchiness between subjects.

To get around the noisiness created by a diverse set of samples, we
built a Random Forest model to identify microbes specific to each side.
For each comparison we identified top X OTUs that were strongly
predictive of one site or another. Generally, OTUs identified in each
location were consistent with known physiological gradients along the
gut axis ((5)). For instance, the proximal mucosa harbored
mucosa-associated facultative anaerobes such as \emph{Actinomyces} and
\emph{Enterobacteraceae} and the aerobic \emph{Psuedomonas} consistent
with the highest oxygen regions of the colon. The distal mucosa was far
more likely to host strictly anaerobic species such as
\emph{Porphyromonas}, \emph{Anaerococcus}, \emph{Finegoldia} and
\emph{Peptoniphilus}. The model was less effective at classifying the
proximal and distal lumenal contents, probably because the samples are
arguably composed of the same material.

We detected \emph{F. nucleatum} and \emph{P. asacharolytica} in 8 and 5
of our subjects, respectively. These bacteria have been shown to be
predictive of colorectal cancer in humans ((9)) and have oncogenic
properties in cell culture and in mice ((17)). Interestingly, while
\emph{F. nucleatum} was found on both sides of the colon, \emph{P.
asacharolytica} was only detected in the distal mucosa. Not much is
known about the distribution of \emph{P. asacharolytica} but given its
documented anaerobic characteristics and asacharolytic metabolism, it
might not be surprising that it resides in the less-oygen-rich and
proteinaceous distal mucosa ((4)). In studies examining bacteria on
colorectal cancer tumors, \emph{F. nucleatum} is more commonly detected
on proximal-sided tumors, and distribution of \emph{F. nucleatum}
decreases along the colon to rectum ((18)). Of the (8) (40\%)
individuals positive for \emph{F. nucleatum} in this study, the bacteria
was spread across the proximal mucosa, distal lumen and distal mucosa.
The presence of \emph{F. nucleatum} in a healthy individual is not
necessarily linked to the development of future colorectal cancers.
Because of the spatial distribution of the \emph{F. nucleatum} in our
sample set, we cannot develop a model for the role of \emph{F.
nucleatum} in the healthy colon. Data examining bacterial biofilms on
CRC tumors suggests that \emph{Fusobacteria} species are more commonly
found both on proximal tumors and in biofilms, indicating that it is not
only the presence of the bacteria but the organization of the tumor
community that contributes to \emph{Fusobacterium's} role in
tumorigenesis ((8)). Finally, \emph{Fusobacteria} and
\emph{Porphyromonas} species have been known to not only co-occur on CRC
tumors but also to synergistically promote tumorigenesis in an oral
cancer model ((19), (20)). Thus, further analysis of the distribution
and activities of these pathogens may elucidate a mechanism for
development of CRC tumors in the proximal or distal colon.

Specific comparisons of our findings to previously published gut
biogeography studies are additionally confounded by the use of bowel
preparation methods in most other studies. A rare report of a
matched-colonoscopy study that sampled 18 patient's colonic mucosa and
lumenal contents prior to and after bowel cleansing ((21)). This group
found that mucosa and lumenal samples were distinguishable prior to
bowel cleansing, but that bowel preparation resulted in an increase in
shared OTUs between each site ((21)). Bowel cleansing not only made the
samples harder to distinguish, it resulted in decreases in diversity
across sites. Further, the differences were not great enough to overcome
interpersonal differences between subjects. So, bowel preparation
clearly induces bias into the microbes recovered from sampling the lumen
or mucosa of a prepared bowel. Thus our findings of specific bacteria at
each site of the colon are strengthened by the lack of bowel
preparation.

Microbiome-based diagnostics are increasingly being explored as
non-invasive tools to survey for the development of colon cancer. Random
Forest models have been used by our group and others to increase the
detection sensitivity of CRC tumors. Indeed, our group found that a
classification model that used microbiome data in combination with Fecal
Immunohistochemical Test (FIT) results could correctly identify both
carcinoma and adenoma lesions from communities of stool at exit and it
performed much better than FIT alone. Further work from our lab has
shown that microbiome profiling of the FIT cartridge contents
sufficiently represented the stool community ((9)). One caveat of the
FIT study was that there was not sufficient information to test if a
classification model could differentiate between proximal and distal CRC
tumors based on exit stool sample alone, but we would hypothesize that
would not be effective. Given that our results showed that the stool
most accurately reflects the community of the distal lumen, we likely
cannot use Random Forest of stool samples to diagnose any changes in the
proximal or mucosal communities.

By revealing specific differences in microbial populations at each
location in the gut via sampling an unprepared bowel, we can begin to
form hypothesies about how specific host-microbe interactions can affect
oncogenesis of proximal and distal CRC tumors. To this point, 16S
community profiling studies do not provide enough information to probe
these questions. Our sample set of matched proximal, distal, lumenal and
mucosal samples from colons that have not undergone bowel preparation
presents a unique opportunity to explore further questions about the
microbiome along the gut axis. Specifically, examining metagenomic,
metabolomic and host interactions at each site will be useful in further
characterizing the host-microbe interactions in the development of
proximal and distal colorectal cancer.

@@

At a baseline visit, subjects were consented and given a home collection
stool kit (Source of kit supplies). At least one week prior to the
scheduled colonoscopy, subjects were to collect whole stool at home and
two swabs of a Fecal Immunohistochemical Test cartridge (Polymedco Inc.)
and ship the samples to the University. Notably, subjects did not
undergo any bowel preparation method prior to sampling. On procedure
day, subjects reported to the Michigan Clinical Research Unit at the
University of Michigan Health System. Patients were consciously sedated
using Fentanyl, Versed and/or Benadryl as appropriate. A flexible
sigmoidoscope was first inserted about 25cm into the colon and endoscopy
brush used to collect lumenal/stool contents. Two lumenal samples were
collected and the contents immediately deposited into RNAlater (source)
and flash-frozen in liquid nitrogen. The brushes were withdrawn and
biopsy forceps were used to collect mucosal biopsies on sections of the
colon that were pink and free of stool matter. Three mucosal biopsies
were collected and flash-frozen in RNAlater. These samples comprised the
distal or distal colon samples. The sigmoidoscope was then withdrawn and
a pediatric colonoscope was inserted to reach the ascending colon.
Samples were then collected as in the distal colon and the colonoscope
withdrawn. All samples were stored at -80 C until study completion.

@@

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Unprepared colonoscopy identifies microbiota specific to the proximal and distal human colon},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{Unprepared colonoscopy identifies microbiota specific to the proximal
and distal human colon}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

@@

Running title: Unprepared colonoscopy identifies proximal and distal
human colonic microbiota

@@

The human colon contains chemicals and nutrients that change along the
proximal to distal gut axis. These gradients create microenvironments
that affect the distrubution and composition of the gut microbiota. The
microbiome has been implicated in the colonic diseases colorectal cancer
(CRC) and inflammatory bowel disease (IBD). Further, these diseases
exhibit different symptoms depending on the location of the colon they
are found in. CRC tumors of the proximal and distal colon are
morphologically and genetically distinct. Similarly, bowel diseases such
as Crohn's are typically exacerbated in the proximal instestine while
ulcerative colitis patients often experience symptoms in the distal
colon. Previous analysis of the fecal microbiota from healthy and CRC or
IBD patients has revealed different microbial signatures associated with
these diseases. We extended these observations of the fecal microbiome
to include analysis of the proximal and distal healthy human colon. We
used a two-colonoscope approach on subjects that had not undergone
standard bowel preparation procedure. This technique allowed us to
characterize the native proximal and distal luminal and mucosal
microbiome without prior chemical disruption. 16S rRNA gene sequencing
was performed on proximal and distal mucosal and luminal biopsies and
home-collected stool for 20 healthy individuals. Diversity analysis
revealed that each site contained a diverse community, and that a
patient's samples were more similar to each other than to that of other
individuals. Comparison of all samples to fecal samples taken at exit
uncovered that the feces were most similar to samples taken from the
distal lumen, likely reflecting the anatomical structure of the colon.
Since we could not differentiate sites along the colon based on
community structure or community membership alone, we employed the
Random Forest machine-learning algorithm to identify key species that
distinguish biogeographical sites. Random Forest classification models
were built using taxa abundance and sample location and revealed
distinct populations that were found in each location.
\emph{Peptoniphilus, Anaerococcus, Enterobacteraceae, Pseudomonas} and
\emph{Actinomyces} were most likely to be found in mucosal samples
versus luminal samples (AUC = 0.925). The classification model performed
well (AUC = 0.912) when classifying mucosal samples into proximal or
distal sides, but separating luminal samples from each side proved more
challenging (AUC = 0.755). The distal mucosa was found to have high
populations of \emph{Finegoldia, Murdochiella} and \emph{Porphyromonas}.
Proximal and distal luminal samples were comprised of many of the same
taxa, likely reflecting the fact that stool moves along the colon from
the proximal to distal end. By sampling the unprepped human colon, our
results have identified distinct bacterial populations native to the
proximal and distal sides. Further investigation of these bacteria may
elucidate if and how these groups contribute to different pathogenesis
processes on the respective sides of the colon.

@@

The human colon is an ecosystem comprised of several different
microenvironments inhabited by resident bacterial members of the
microbiome. Concentrations of oxygen, water and anti-microbial peptides
change along the gut axis and influence what populations of microbes
reside in each location. Microenvironments differ not only
longitutinally along the colon, but latitudinally from the epithelium to
mucosa to intestinal lumen, offering several sites for different
microbial communities to flourish. The identity of these specific
microbes and communities are important for understanding the etiology of
complex colon diseases such as Inflammatory Bowel Disease (IBD) and
Colorectal Cancer (CRC). IBD and CRC are known to be preceded or
accelerated by perturbations in gut microbes ((1), (2), (3)). The
severity, symptoms, morbidity and mortality of these diseases is known
to vary based upon the biogeographical location in which they occur. For
instance, CRC tumors that arise on the distal side of the colon are
infiltrating lesions that present with painful symptoms ((4)). In
contrast, 47\% of CRCs are caused by proximal-sided colon tumors that
are sessile and form along the wall of the colon, often remaining
asymptomatic until advanced carcinogenesis ((4)). The distal and
proximal sides of the colon differ in the amount of inflammation present
and the genomic instability of precancerous cells, respectively, in
addition to variation in the previously mentioned chemical gradients
((1), (5), (6)). In IBD patients, disease flares in the distal colon are
usually indicative of ulcerative colitis (UC), whereas Crohn's disease
(CD) patients typically experience disease in the small intestine, ileum
and proximal colon ((2)). UC presents as large and highly inflammed
mucosal ulcers, where as CD lesions are often smaller and have areas of
normal tissue distributed amongst the flare ((2)). Thus, given the
varied physiology of the proximal-distal axis of the colon and known
differences in disease patterns at these sites, symbiotic microbes and
their metabolites likely vary as well, and may influence the
heterogenous disease prognoses of IBD and CRC. Because CRC is a
long-term complication of IBD, the distribution of microbes is important
to understanding the pathophysiology of both diseases.

Several recent findings have shown that development and progression of
IBD or CRC can be attributed to specific molecular events as a result of
interactions between the gut microbiota and human host ((1), (3)). For
instance, comparison of the bacteria present on CRC tumors with those
found on nearby healthy tissue has identified specific species that are
tumor-associated ((7)). These species include the oral pathogens
\emph{Fusobacterium nucleatum} and \emph{Porphyromonas asacharolytica}.
Interestingly, these periodontal pathogens have been highly predictive
of whether a patient had CRC tumors or not in our human stool
classification studies ((8)). \emph{F. nucleatum} has also been found to
be elevated in the stool and biopsies of patients with IBD as compared
to healthy controls (Strauss2011). Furthermore, studies of \emph{F.
nucleatum} isolated from mucosal biopsies showed that more invasive
\emph{F. nucleatum} positively correlates with IBD disease level
(Strauss2011). Like many intestinal pathogens, the bacteria appear to
have a high-impact despite being lowly-abundant in the community ((2)).
The molecular capabilities of these rare taxa may contribute to the
colonic disease state. These studies often examined only shed human
stool or the small intestine, preventing fine-resolution analysis of
paired samples from the proximal and distal sides of the colon.
Similarly, comparisons of on- or off-tumor/lesion bacteria rarely have
matched tissue from the other side of the colon from the same patient,
limiting what conclusions can be drawn about the colonic microbiome
overall, let alone at that specific site. Due to these limitations, the
contribution of the gut microbiota to IBD and CRC disease location in
the colon is largely undefined. Characterizing these communities could
provide needed insight into disease etiology, including how the
disruption of the healthy community could promote the initiation or
proliferation of the distinct proximal and distal CRC tumors or IBD
flares.

The few existing profiles of the microbial biogeography of the gut have
been limited by sample collection methods. The majority of human gut
microbiome studies have been performed on whole shed stool or on samples
collected during colonoscopy procedures. While the latter method allows
investigators to acquire samples from inside the human colon, typically
this procedure is preceded by the use of bowel preparation methods such
as the consumption of laxatives to cleanse the bowel. Bowel preparation
is essential for detecting cancerous or precancerous lesions in the
colon, but complicates microbiome profiling as the chemicals strip the
bowel of contents and disrupt the mucosal layer ((9), (10)). As such,
what little information we do have about the biogeographical
distribution of the microbes in the proximal and distal colon is
confounded by the bowel preparation procedure.

Here we aimed to address the limitations of previous studies and
effectively identify the microbes specific to the lumen and mucosa of
the proximal and distal healthy human colon. Our design used an
unprepared colonoscopy technique to sample the natural community of each
location of the gut without prior disruption of the native bacteria in
20 healthy volunteers. To address the inherent inter-individual
variation in human microbiomes, we used a machine-learning
classification algorithm trained on curated 16S rRNA sequencing reads to
identify microbes specific to each location. We found that our
classification models were able to separate mucosal and lumenal samples
as well as differentiate between sides of the colon based on populations
of specific microbes. By identifying the specific microbes we are poised
to ask if and how the presence or disruption of the microbes at each
site contribute to the development of the specific tumor subtypes of CRC
in the proximal and distal human colon.

@@

Lumenal and mucosal samples were collected from the proximal and distal
colon of 20 healthy humans that had not undergone bowel preparation
(Figure 1). Participants also collected stool at home one week prior to
the procedure. To characterize the bacterial communities present at
these sites, 16S rRNA gene sequencing was performed on extracted DNA
from each sample. Each site was primarily dominated by \emph{Firmicutes}
and \emph{Bacteriodetes} (Figure 2A), consistent with known variability
in human microbiome research ((11)). Likewise, samples had varying
levels of diversity at each site, irrespective of the individual (Figure
2B). For example, the proximal mucosa was more diverse than the distal
for some individuals while the opposite was true for others. Therefore
we could not identify a clear pattern of changes in microbial diversity
along the gut axis.

@@

Here we identified bacterial taxa that were specific to the lumen and
mucosa of the proximal and distal sides of the human colon from samples
collected during unprepared colonoscopy. We found that all locations
contained a range of phyla and a range of diversity, but that there was
a wide variability between subjects. Pairwise comparisons of each of the
sites revealed that the proximal mucosa and lumen were most similar to
each other. Further, comparison of colonoscopy-collected samples with
samples collected from stool at home showed that the distal lumen is
most similar to stool at exit. Random Forest models built on OTU
relative abundances from each sample identified microbes that are
particular to each location of the colon. Finally, we were able to
detect some bacterial OTUs associated with colonic disease in our
healthy patient cohort. Using unprepped colonoscopy and machine
learning, we have identified bacterial phyla specific to the healthy
proximal and distal human colon.

When examining the relative abundance of the different phyla at each
site, there was a wide amount of variation for each phyla with
communities primarily dominated by the \emph{Bacteriodes} and
\emph{Firmicutes}. This likely reflects not only the variability between
human subjects, casued by differences in age, gender, diet, but also
reports of biogeogrphical ``patchiness'' in the gut microbiome. Several
studies have noted that the bacteria recoverable from the same mucosal
sample location can be vastly different when the samples are taken just
1 cm away from each other (15). Similar patchiness is also observed in
lumenal contents and fecal samples themselves; there is observed
separation of different interacting microbes along the length of a stool
sample, for instance (16). That said, across our samples the mucosal
samples harbor more \emph{Proteobacteria}, consistent with previous
studies comparing mucosal swabs to lumenal content in humans (5). Hence,
the conclusions we can draw from phyla analysis are likely impacted by
patchiness between subjects.

To get around the noisiness from a diverse set of samples, we built a
Random Forest model to identify microbes specific to each side. For each
comparison we identified top X OTUs that were strongly predictive of one
site or another. Generally, OTUs identified in each location were
consistent with known physiological gradients along the gut axis (6).
For instance, the proximal mucosa contains the highest oxygen
concentratons of the colon and harbored mucosa-associated facultative
anaerobes such as \emph{Actinomyces} and \emph{Enterobacteraceae} and
aerobic \emph{Psuedomonas}. The distal mucosa was far more likely to
host strictly anaerobic species such as \emph{Porphyromonas},
\emph{Anaerococcus}, \emph{Finegoldia} and \emph{Peptoniphilus}. The
model was less effective at classifying the proximal and distal lumenal
contents, probably because the samples are arguably composed of the same
bacteria but differ in water content.

We detected \emph{F. nucleatum} and \emph{P. asacharolytica} in 8 and 5
of our subjects, respectively. These bacteria have been shown to be
predictive of colorectal cancer in humans (8) and have oncogenic
properties in cell culture and in mice (17). Interestingly, while
\emph{F. nucleatum} was found on both sides of the colon, \emph{P.
asacharolytica} was only detected in the distal mucosa. Not much is
known about the distribution of \emph{P. asacharolytica} but given its
documented anaerobic characteristics and asacharolytic metabolism, it
might not be surprising that it resides in the less-oygen-rich and
proteinaceous distal mucosa ((5)). In studies examining bacteria on
colorectal cancer tumors, \emph{F. nucleatum} is more commonly detected
on proximal-sided tumors, and distribution of \emph{F. nucleatum}
decreases along the colon to rectum ((18)). Of the 8 (40\%) individuals
positive for \emph{F. nucleatum} in this study, the bacterium was spread
across the proximal mucosa, distal lumen and distal mucosa. The presence
of \emph{F. nucleatum} in a healthy individual is not necessarily linked
to the development of future colorectal cancers (cite). Data examining
bacterial biofilms on CRC tumors suggests that \emph{Fusobacteria}
species are more commonly found on proximal tumors and in biofilms,
indicating that it is not only the presence of the bacteria but the
organization of the tumor community that contributes to
\emph{Fusobacterium's} role in tumorigenesis ((7)). Finally,
\emph{Fusobacterium} and \emph{Porphyromonas} species have been known to
not only co-occur on CRC tumors but also to synergistically promote
tumorigenesis in an oral cancer model ((19), (20)). Thus, further
analysis of the distribution and activities of these pathogens may
elucidate a mechanism for development of IBD or CRC subtypes in the
proximal or distal colon.

The \emph{Fusobacterium} species \emph{nucleatum} and \emph{varium} have
been commonly isolated from mucosal biopsies of patients with IBD
(cite). Laboratory experiments with these isolates have shown that
disease-isolated \emph{F. nucleatum} are more invasive and stimulate
more TNF-a production than strains from healthy individuals ((21)),
suggesting the bacteria may increase inflammation in the gut as well
(cite - Dharmani 2011). \emph{F. varium} isolated from UC patients
caused colonic ulcers in an experimental mouse model (Ohkusa 2003).
\emph{F. varium} was only detected in 3 of our study participants and
two of those samples were isolated form the proximal mucosa (cite
figure). Although there is low occurrance in our study, \emph{F. varium}
is most commonly isolated from UC patient biopsies from the ileum or
cecum (cite), suggesting this species may exhibit preference for the
different environmental conditions of these gastrointestinal sites.
Further work will assess how gut environment may select for species
which then cause localized disease.

Specific comparisons of our findings to previously published gut
biogeography studies are additionally confounded by the use of bowel
preparation methods in most other studies. A rare report of a
matched-colonoscopy study sampled 18 patient's colonic mucosa and
lumenal contents prior to and after bowel cleansing ((22)). This group
found that mucosal and lumenal samples were distinguishable prior to
bowel cleansing, but that bowel preparation resulted in an increase in
shared OTUs between each site ((22)). Bowel cleansing not only made the
samples harder to distinguish, it resulted in decreases in diversity
across sites. Further, the differences were not great enough to overcome
interpersonal differences between subjects. Bowel preparation clearly
induces bias into the microbes recovered from sampling the lumen or
mucosa of a prepared bowel. Thus our findings in this study are
strengthened by the lack of bowel preparation.

By revealing specific differences in microbial populations at each
location in the gut via sampling an unprepared bowel, we can begin to
form hypothesies about how specific host-microbe interactions can affect
disease progression of proximal and distal CRC and IBD subtypes. To this
point, 16S rRNA gene sequecing community profiling studies do not
provide enough information to fully probe these questions. In
particular, 16S sequencing cannot not profile the host characteristics
at each site. Combining the unprepared colonoscopy approach with
analysis of multi-omic sequencing data may be useful in further
characterizing host-microbiome interactions along the gut axis for both
health and disease.

@@

At a baseline visit, subjects were consented and given a home collection
stool kit (Source of kit supplies). At least one week prior to the
scheduled colonoscopy, subjects were to collect whole stool at home and
ship the samples to the University on ice. Notably, subjects did not
undergo any bowel preparation method prior to sampling. On the procedure
day, subjects reported to the Michigan Clinical Research Unit at the
University of Michigan Health System. Patients were consciously sedated
using Fentanyl, Versed and/or Benadryl as appropriate. A flexible
sigmoidoscope was first inserted about 25cm into the colon and endoscopy
brush used to collect lumenal/stool contents. Two lumenal samples were
collected and the contents immediately deposited into RNAlater (source)
and flash-frozen in liquid nitrogen. The brushes were withdrawn and
biopsy forceps were used to collect mucosal biopsies on sections of the
colon that were pink and free of stool matter. Three mucosal biopsies
were collected and flash-frozen in RNAlater. These samples comprised the
distal or distal colon samples. The sigmoidoscope was then withdrawn and
a pediatric colonoscope was inserted to reach the ascending colon.
Samples were then collected as in the distal colon and the colonoscope
withdrawn. All samples were stored at -80 C until study completion.

@@

\paragraph{Data availability}\label{data-availability}
